---
title: "Assignment 2"
output: pdf_document
date: "2022-09-09"
---

# Part 1

## Red Wine Quality

a.) Reads winequality-red.csv into 'redwine'.

```{r}
redwine = read.csv("winequality-red.csv", sep = ",", header = TRUE)

summary(redwine)
```

b.) Determines the median of red wine quality and mean of alcohol taste score. The median is 6 and the mean is 10.42298.

```{r}
median(redwine$quality)

mean(redwine$alcohol)
```

c.) Plots a scatter plot of pH to alcohol quality.

```{r}
library(ggplot2)
ggplot(redwine, aes(x = pH, y = quality)) + 
  geom_point() +  
  labs(x = "pH Level", y = "Alcohol Quality", title = "pH Level vs. Alcohol Quality") +
  theme(plot.title = element_text(hjust = 0.5))
```

d.) Creates a new column in the data frame 'redwine' called 'ALevel' and populates it with either 'High' or 'Medium', depending on the alcohol level.

```{r}
redwine$ALevel = " "
for (i in redwine$alcohol)
{
  if (i > 10.2)
    redwine$ALevel[which(redwine$alcohol == i)] = "High"
  else
    redwine$ALevel[which(redwine$alcohol == i)] = "Medium"
}

str(redwine)
```

Determines the sulphate to chloride ratio, as well as converting 'ALevel' into a factor, before plotting the two variables on a box plot.

```{r}
sulphate_to_chloride = redwine$sulphates / redwine$chlorides
redwine$ALevel = as.factor(redwine$ALevel)
ggplot(redwine, aes(x = ALevel, y = sulphate_to_chloride)) + 
  geom_boxplot() + 
  labs(y = "Sulphate to Chloride Ratio", title = "Sulphate to Chloride Ratio Boxplot") + 
  theme(plot.title = element_text(hjust = 0.5))
```

Determines the amount of red wine who's alcohol level is considered as 'High' or exceeding 10.2. The answer is 757.

```{r}
sum(redwine$ALevel == "High")
```

e.) A histogram of 'total_sulfur_dioxide' for both 'High' and 'Medium' 'ALevel' labels.

```{r}
ggplot(redwine, aes(x = total_sulfur_dioxide)) +
  geom_histogram() + 
  labs(x = "Total Sulfur Dioxide", y = "Count", title = "Total Sulfur Dioxide Histogram") +
  theme(plot.title = element_text(hjust = 0.5))
```

f.) Scatter plot of pH vs. alcohol level.

```{r}
ggplot(redwine, aes(x = pH, y = alcohol)) + 
  geom_point() +
  labs(x = "pH Level", y = "Alcohol Level", title = "pH Level vs. Acohol Level") +
  theme(plot.title = element_text(hjust = 0.5))
```

Scatter plot of residual sugar vs. alcohol level.

```{r}
ggplot(redwine, aes(x = residual_sugar, y = alcohol)) + 
  geom_point() +
  labs(x = "Residual Sugars", y = "Alcohol Level", title = "Residual Sugars vs. Alcohol Level") +
  theme(plot.title = element_text(hjust = 0.5))
```

From the two plots, it can be hypothesized that there are much more people who prefer acidic, less alcoholic and less sweet wine. According to the first plot, there is an abundant amount of people who prefer an acidic wine which is less alcoholic, and based on the second plot, more people preferred less sweetened wine regardless of whether they preferred strong or weaker alcohol.

------------------------------------------------------------------------

# Part 2

## Forest Fires

The file, 'forestfires.csv' is read into 'forest_fires', and the column 'DC' in the data frame is converted into integer.

```{r}
forest_fires = read.csv("forestfires.csv", sep = ",", header = TRUE)
forest_fires$DC = as.integer(forest_fires$DC)
summary(forest_fires)
```

a.) X is Quantitative, Y is Quantitative, month is Qualitative, day is Qualitative, FFMC is Quantitative, DMC is Quantitative, DC is Quantitative, ISI is Quantitative, temp is Quantitative, RH is Quantitative, wind is Quantitative, rain is Quantitative, area is Quantitative.

b.) Determines the ranges, means and standard deviations for each column of 'forest_fires', and a data frame for all three were made using the data calculated.

```{r}
predictor_r = c("X min", "X max", "Y min", "Y max", "FFMC min", "FFMC max", "DMC min", "DMC max", "DC min", "DC max",
                "ISI min", "ISI max", "temp min", "temp max", "RH min", "RH max", "wind min", "wind max", "rain min", 
                "rain max", "area min", "area max")

range = c(range(forest_fires$X), range(forest_fires$Y), range(forest_fires$FFMC), range(forest_fires$DMC), 
          range(forest_fires$DC), range(forest_fires$ISI), range(forest_fires$temp), range(forest_fires$RH),
          range(forest_fires$wind), range(forest_fires$rain), range(forest_fires$area))
ranges = data.frame(predictor_r, range)


predictor = c("X", "Y", "FFMC", "DMC", "DC", "ISI", "temp", "RH", "wind", "rain", "area")

mean = c(mean(forest_fires$X), mean(forest_fires$Y), mean(forest_fires$FFMC), mean(forest_fires$DMC), 
         mean(forest_fires$DC), mean(forest_fires$ISI), mean(forest_fires$temp), mean(forest_fires$RH),
         mean(forest_fires$wind), mean(forest_fires$rain), mean(forest_fires$area))
means = data.frame(predictor, mean)

standard_deviation = c(sd(forest_fires$X), sd(forest_fires$Y), sd(forest_fires$FFMC), sd(forest_fires$DMC), 
                       sd(forest_fires$DC), sd(forest_fires$ISI), sd(forest_fires$temp), sd(forest_fires$RH),
                       sd(forest_fires$wind), sd(forest_fires$rain), sd(forest_fires$area))
stand_devs = data.frame(predictor, standard_deviation)

str(ranges)
str(means)
str(stand_devs)
```

Determines the most frequency of days of the week in which a forest fire occurred and creates a vector 'week'. And as the vector is sorted from Monday to Sunday, Sunday has the highest frequency.

```{r}
week = c(sum(forest_fires$day == "mon"), sum(forest_fires$day == "tue"), sum(forest_fires$day == "wed"),
         sum(forest_fires$day == "thu"), sum(forest_fires$day == "fri"), sum(forest_fires$day == "sat"),
         sum(forest_fires$day == "sun"))

week
```

c.) Determines the same thing as in b.), however not putting into account rows 40-80 of the 'forest_fire' data frame.

```{r}
new_range = c(range(forest_fires$X[c(1:39, 81:517)]), range(forest_fires$Y[c(1:39, 81:517)]), 
              range(forest_fires$FFMC[c(1:39, 81:517)]), range(forest_fires$DMC[c(1:39, 81:517)]), 
              range(forest_fires$DC[c(1:39, 81:517)]), range(forest_fires$ISI[c(1:39, 81:517)]), 
              range(forest_fires$temp[c(1:39, 81:517)]), range(forest_fires$RH[c(1:39, 81:517)]),
              range(forest_fires$wind[c(1:39, 81:517)]), range(forest_fires$rain[c(1:39, 81:517)]), 
              range(forest_fires$area[c(1:39, 81:517)]))
new_ranges = data.frame(predictor_r, new_range)

new_mean = c(mean(forest_fires$X[c(1:39, 81:517)]), mean(forest_fires$Y[c(1:39, 81:517)]), 
             mean(forest_fires$FFMC[c(1:39, 81:517)]), mean(forest_fires$DMC[c(1:39, 81:517)]), 
             mean(forest_fires$DC[c(1:39, 81:517)]), mean(forest_fires$ISI[c(1:39, 81:517)]), 
             mean(forest_fires$temp[c(1:39, 81:517)]), mean(forest_fires$RH[c(1:39, 81:517)]),
             mean(forest_fires$wind[c(1:39, 81:517)]), mean(forest_fires$rain[c(1:39, 81:517)]), 
             mean(forest_fires$area[c(1:39, 81:517)]))
new_means = data.frame(predictor, new_mean)

new_standard_deviation = c(sd(forest_fires$X[c(1:39, 81:517)]), sd(forest_fires$Y[c(1:39, 81:517)]), 
                           sd(forest_fires$FFMC[c(1:39, 81:517)]), sd(forest_fires$DMC[c(1:39, 81:517)]), 
                           sd(forest_fires$DC[c(1:39, 81:517)]), sd(forest_fires$ISI[c(1:39, 81:517)]), 
                           sd(forest_fires$temp[c(1:39, 81:517)]), sd(forest_fires$RH[c(1:39, 81:517)]),
                           sd(forest_fires$wind[c(1:39, 81:517)]), sd(forest_fires$rain[c(1:39, 81:517)]), 
                           sd(forest_fires$area[c(1:39, 81:517)]))
new_stand_devs = data.frame(predictor, new_standard_deviation)

str(new_ranges)
str(new_means)
str(new_stand_devs)
```

d.) Generates a bar plot showing the frequency of forest forest in each month with a wind speed of greater than 4 (wind \> 4). As such, the plot shows that the month August has the most forest fires with wind \> 4.

```{r}
filter_data = forest_fires[(forest_fires$wind > 4), ]$month
forest_fires_new = data.frame(filter_data)

ggplot(forest_fires_new, aes(x = filter_data)) + 
  geom_bar() + 
  labs(x = "Month", y = "Count", title = "Forest Fire Frequency (Wind > 4)") +
  theme(plot.title = element_text(hjust = 0.5))
```

e.) By plotting on a scatter plot, the Duff Moisture Code Index (DMC) with the Initial Spread Index (ISI), a correlation matrix with these two variables was made.

```{r}
ggplot(forest_fires, aes(x = DMC, y = ISI)) + 
  geom_point() + 
  labs(x = "DMC", y = "ISI", title = "DMC vs. ISI") +
  geom_smooth(method = 'lm') +
  theme(plot.title = element_text(hjust = 0.5))

correlation_matrix = matrix(0, nrow = length(forest_fires$DMC))
correlation_matrix[,1] = correlation_matrix[,1] + 1:517
correlation_matrix = cbind(correlation_matrix, forest_fires$DMC)
correlation_matrix = cbind(correlation_matrix, forest_fires$ISI)
colnames(correlation_matrix) = c("No.", "DMC", "ISI")

str(correlation_matrix)
```

f.)

In order to predict wind speed based on the other variables, pairing it with FFMC, DMC, ISI,temp, and/or RH would be useful. This is because creating a plot with both wind speed and one other variables shows a set of points all relating to a wind speed staying constant regardless of the other variable at random. In turn, this allows us to better understand the wind speed as it may stay constant for several points of the other variable, or it might change.
